The Numerical Accuracy of Computing Electrostatic Interaction in the Inhomogeneous and Correlated Molecular Systems: for the Ewald Summation, SPME and Staggered Mesh Ewald Methods.

### 2012-07-13 longRange.hetero/alltext/publish/resubmit/main.tex
fix bugs according to jctc template.

\documentclass[aps,pre,preprint,unsortedaddress]{revtex4}
% \documentclass[aps,jcp,unsortedaddress]{revtex4-1}
%\documentclass[aps,pre,twocolumn]{revtex4-1}
%\documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvips]{graphicx}
\usepackage{color}
\usepackage{tabularx}

@@

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\renewcommand{\v}[1]{\textbf{\textit{#1}}}
\renewcommand{\d}[1]{\textsf{#1}}
% \allowdisplaybreaks

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

@@

@@

\title{The Numerical Accuracy of Computing Electrostatic Interaction
  in the Inhomogeneous and Correlated Molecular Systems:
  for the Ewald Summation, SPME and Staggered Mesh Ewald Methods.
}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Pingwen Zhang}
\affiliation{LMAM and School of Mathematical Sciences, Peking University}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

@@

The aforementioned fast algorithms have achieved great success
over the past three decades.
However, the fact that the user should provide six working parameters
for these methods may be problematic: it is difficult to find the
most efficient parameter set in such a 6-dimensional parameter space.
% The experiential choices may be either less accurate or wasting
% computational effort.
% This problem can be solved by an
\recheck{
  This problem can be solved by repeating and comparing
  the force calculation by
  using different parameter sets
  on a representative snapshot of the system,
  see e.g. Ref.~\cite{abraham2011optimization}.}
\recheck{An other approach is the 
  in-depth theoretical study}
of the error introduced by the algorithms, namely the
error estimate~\cite{kolafa1992cutoff, hummer1995numerical,
  petersen1995accuracy, deserno1998mue2, stern2008mesh,
  wang2010optimizing, neelov2010interlaced, ballenegger2012convert},
by which the error is described by a function of the working
parameters.  Using the error estimates, a work flow has been designed to
automatically determine the nearly optimal combination of parameters,
in the sense that it is the fastest one satisfying a demanding
accuracy~\cite{wang2010optimizing}. However, so far, all error
estimates assume the homogeneity of the charge distribution, and the
independency of any pair of charges (i.e. the charges are
uncorrelated), but neither of the assumptions is satisfactorily
fulfilled in most molecular
systems of practical interest.  For example, the lipid bilayer membrane
solving in water: the hydrophobic tail(s) of the lipids carry no charge, and
the charges in the system
are correlated by the covalent bonds, the van der Waals
interaction, the hydrogen bonding network, etc..  Therefore, the current
error estimates may be problematic, or even not applicable.  Several
papers have shown that due to the correlation of charges, the error estimates
obviously deviate from the real
error~\cite{deserno1998mue2,wang2010optimizing,ballenegger2012convert}.

The purpose of this manuscript is to develop a reliable  error
estimate in inhomogeneous and correlated molecular systems.
We estimate the force error because the application is the molecular
dynamics (MD) simulation.  Firstly, a general error estimate
framework for the pairwise interaction is set up.
\recheck{We show that the
  error may be approximately
  decomposed into three additive parts: the homogeneity error,
the inhomogeneity error and the correlation error~\cite{wang2012}.}
% These error estimates are only of theoretical interest, and
% not applicable, because the computational costs of the error estimates
% are $\mathcal O(N^2)$ (homogeneity and inhomogeneity error)
% and $\mathcal O(N^3)$ (correlation error).
% If the error force has a special form so that the error estimates
% can be written as convolutions, the computational costs
% are reduced to $\mathcal O(N\log N)$ and 
% $\mathcal O(N^2\log N)$, respectively.
The computational cost of the homogeneity and inhomogeneity
error is $\mathcal O(N\log N)$ when the error force kernel is
of convolution form.
The  correlation error can also be
estimated at a cost of $\mathcal O(N\log N)$
by the nearest neighbor approximation
technique, which \redc{assumes} the error  is dominated by  short-range correlation.
This
is the case for most the liquid and gas systems far from critical point.
Generally speaking,
when the error force can be written as a pairwise form, 
all estimate are derived easily under this framework.
Following this framework, we
develop the error estimates for the Ewald summation, the SPME 
and the staggered method Ewald methods.   The estimates
of the homogeneity and inhomogeneity errors are tested by two
randomly distributed inhomogeneous charges systems,
% (i.e. the charges are not correlated),
and is proved to
be very sharp.
The error estimates are also tested in a real inhomogeneous water system,
in which the charge correlation plays a very important role
on the force error. By 
including the nearest neighbor approximation of
\redc{the} correlation error, the quality of the estimates
is impressively enhanced.
This paper is ended by conclusions and remarks in
section~\ref{sec:conclusions}.

@@

Here are some remarks concerning Theorem~\ref{thm:tmp1} and Corollary~\ref{thm:tmp2}:

@@

  Eqn.~\eqref{eqn:def-rhoq} and \eqref{eqn:def-rhoq2} are called \emph{the

@@

  $\rho^{(2)}(\v r, \v r')$ in Eqn.~\eqref{eqn:def-rho2}

@@

  Eqn.~\eqref{eqn:def-c} is the \emph{pairwise charge correlation function},

@@

  Also see Sec.~\ref{sec:example3} for example.

@@

  of the interaction (see Eqn.~\eqref{eqn:thm-error-force}) is the key to derive
  the error estimates Eqn.~\eqref{eqn:thm-meanf} and \eqref{eqn:thm-error}.

@@

  By setting $q_i = q_j = 1$, the corollary~\ref{thm:tmp2} proves

@@

  three terms on the r.h.s. of Eqn.~\eqref{eqn:thm-error} are called the

@@

  The first term on the r.h.s. of \eqref{eqn:error-split} is called

@@

  Theorem~\ref{thm:tmp1} calculates the homogeneity and inhomogeneity

@@

  the corollary~\ref{thm:tmp2},

@@

\item A non-vanishing mean error force~\eqref{eqn:thm-meanf} was proved

@@

Starting from Eqn.~\eqref{eqn:prf-error-step3},

@@

Eqn.~\eqref{eqn:def-rho2}.

@@

second \redc{term} on the r.h.s. of the Eqn.~\eqref{eqn:split-rho-q2} is

@@

on the r.h.s. of Eqn.~\eqref{eqn:split-rho-q2} is actually the

@@

By inserting Eqn.~\eqref{eqn:fnc-correlation-water} into

@@

As an example, we consider a typical three point charge water model
TIP3P~\cite{jorgensen1983comparison}: each hydrogen atom carries a
charge of $q_{\textrm{H}} = +0.417e$ and each oxygen atom carries a
charge of $q_{\textrm{O}} = -0.834e$.  \recheck{ There are two types
  of correlation in this system: (1) the bonded correlation
  that originates from the rigid H-O
  bond and H-O-H angle within one molecule,
  and (2) the
  non-bonded correlation or the
  intermolecular correlation that originates from
  a complex coupling of various intermolecular interaction, such as
  the van der Waals interaction,
  the hydrogen bonding and etc..
% Within one molecule, the H-O bond and
% H-O-H angle are rigid, so the correlation within one molecule is
% considered important.
% Therefore, in this case,
}

@@

neighbor by the atoms within the same water molecule.  Denoting the

@@

The two ensemble averages in Eqn.~\eqref{eqn:c-error-1}

@@

Eqn.~\eqref{eqn:c-error-1} is by the Fourier transform

@@

the Eqn.~\eqref{eqn:c-error-1} becomes:

@@

system, as illustrated by Fig.~\ref{fig:tmp-add0}.

@@

% estimate~\eqref{eqn:c-error-water}

@@

standard long-range algorithms.}

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{correlation.1.eps}
  \caption{The schematic plot of two interacting water molecules.
    When we calculate the force error of the left oxygen atom (it is now
    a testing charge), the
    correlation with its bonded hydrogens are not considered, but
    the correlations within any neighboring molecule are considered
    by the bonded nearest neighbor approximation,
    as indicated by the solid lines with two arrows.}
  \label{fig:tmp-add0}
\end{figure}

\recheck{
  For the non-bonded correlation,
  we define the nearest neighbor
  by non-bonded atoms closer than a certain radius $R_g$ (neighboring range).
  If the system is homogeneous and isotropic, the pair
  correlation function $\rho^{(2)}(\v r', \v r'')$ in
  Eqn.~\eqref{eqn:start-corr} can be represented by the radial
  distribution functions (RDFs).
  Following the same way as the bonded correlation,
  the non-bonded correlation error can be estimated:
  the functions $T_{\textrm{O}}$ and $T_{\textrm{H}}$
  are given by
  \begin{align}\label{eqn:nb-nna-1}
    \hat T_{\textrm{O}} (\v m)
    &= \int_0^{R_g}
    [q_{_{\textrm{H}}} \rho_{_{\textrm{H}}} \tilde g_{_{\textrm{OH}}}(r)
    +q_{_{\textrm{O}}} \rho_{_{\textrm{O}}} \tilde g_{_{\textrm{OO}}}(r)]
    \,\frac{\sin(2\pi m r)}{2\pi mr}\,4\pi r^2 \d dr,\\\label{eqn:nb-nna-2}
    \hat T_{\textrm{H}} (\v m)
    &= \int_0^{R_g}
    [q_{_{\textrm{H}}} \rho_{_{\textrm{H}}} \tilde g_{_{\textrm{HH}}}(r)
    +q_{_{\textrm{O}}} \rho_{_{\textrm{O}}} \tilde g_{_{\textrm{OH}}}(r)]
    \,\frac{\sin(2\pi m r)}{2\pi mr}\,4\pi r^2 \d dr,
  \end{align}
  where
  $\tilde g_{_{\textrm{OO}}}(r) = g_{_{\textrm{OO}}}(r) - 1$,
  $\tilde g_{_{\textrm{OH}}}(r) = g_{_{\textrm{OH}}}(r) - 1$ and
  $\tilde g_{_{\textrm{HH}}}(r) = g_{_{\textrm{HH}}}(r) - 1$.
  $g_{_{\textrm{OO}}}(r)$, $g_{_{\textrm{OH}}}(r)$ and $g_{_{\textrm{HH}}}(r)$
  are the RDFs of oxygen-oxygen, oxygen-hydrogen and hydrogen-hydrogen,
  respectively. $\rho_{_{\textrm{H}}}$ and $\rho_{_{\textrm{O}}}$ are
  the homogeneous number densities
  of the hydrogen and oxygen atoms, respectively.
  For a short-range correlated system, the RDFs converges to 1
  as $r$ goes to infinity, so the integrals in Eqn.~\eqref{eqn:nb-nna-1}
  and \eqref{eqn:nb-nna-2} are convergent with respect to
  increasing $R_g$.

@@

The complementary error function $\textrm{erfc}(r)$ in
Eqn.~\eqref{eqn:tmp2} decays exponentially as $r$ goes to infinity.
Therefore, it is a short-range interaction in the real space, which
can be calculated by the standard cut-off and neighbor list
method~\cite{frenkel02b} at a cost of $\mathcal O(N)$.  The reciprocal
part also decays exponentially as the magnitude of the Fourier mode
$\vert\v m\vert$ increases. Therefore, in practice, the infinite
summation in Eqn.~\eqref{eqn:tmp3} is truncated, and only a finite
summation is calculated:
% Both the direct and the reciprocal parts are calculated by cut-off in
% real and reciprocal space, respectively.
% For the reciprocal part the cut-offed energy reads
\begin{align}\label{eqn:tmp6}
  E^{\textrm{tr}}_{\textrm{rec}} & =
  % \frac1{2\pi V}
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  % \frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m)   
  f(\v m)\, S(\v m) S(-\v m),
\end{align}
where, for short,
\begin{align}
  f(\v m) := \frac{\exp(-\pi^2 \v m^2 / \beta^2)}{2\pi V \v m^2},
\end{align}
and $K_\alpha$ is the number of Fourier modes used on direction
$\alpha$.  The truncated reciprocal force acting on particle $i$ is
\begin{align}
  % \nonumber
  \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
  % &= 
  % q_i 
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_\alpha\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % \v g(\v m) \,
  % S(-\v m)\,
  % e^{2\pi i\v m\cdot\v r_i}, \\
  \label{eqn:tmp8}
  &= 
  \sum_j   q_iq_j
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r_{ij}},
\end{align}
where
\begin{align}
  \v g(\v m) = -4 \pi \v m i f(\v m).
\end{align}

@@

Both the error force kernel of the direct and reciprocal
part of the Ewald summation can be
written in the form of Eqn.~\eqref{eqn:cor-kernel}.  The direct
part is:
% The direct part is a short-range interaction in real space, so it is
% usually treated by the cut-off method, which gives a error force
% kernel of
\begin{align}
  \v K^{\textrm{cut}}_{\textrm{dir}}(\v r) =
  \left\{
    \begin{aligned}
      &\,0, &\quad & r \leq r_c;\\
      &
      \Big[
      \frac{2\beta}{\sqrt\pi} e^{-\beta^2r^2} + \frac{\textrm{erfc}(\beta r)}{r}
      \Big]\frac{\v r}{r^2}
      ,& & r > r_c,
      % \frac{\textrm{erfc}(\beta\vert\v r\vert)}{\vert\v r\vert}
    \end{aligned}
  \right.
\end{align}
where $r = \vert\v r\vert$, and $r_c$ is the cut-off radius in the
real space.
% The reciprocal error force, namely the difference between
% the the exact and truncated reciprocal force, is
% \begin{align}
%   \Delta \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
%   = & - 
%   \sum_j\,q_iq_j
%   \sum_{\vert m_\alpha\vert \geq K_\alpha/2}
%   \v g(\v m) \,
%   e^{2\pi i\v m\cdot\v r_{ij}}.
% \end{align}
% Therefore,
The reciprocal error force kernel is given by:
\begin{align}
  \v K^{\textrm{tr}}_{\textrm{rec}}(\v r) =
  \sum_{
      \vert m_\alpha\vert \geq K_\alpha/2}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r}.
\end{align}
By the corollary~\ref{thm:tmp2}, the error estimate of the
truncated Ewald summation is straightforward.

% Ref. \cite{short} pointed out that the inhomogeneity error is more
% harmful than the homogeneity error, so it is interesting to analyze
% the inhomogeneity error. Apply the Fourier transform on both sides of
% Eqn. \eqref{eqn:tmp10} and \eqref{eqn:tmp18}:
% \begin{align}  
%   \langle\Delta\v F_{\textrm{dir}}\rangle^\wedge(\v m)
%   &= q_i\,\hat{\v K}^c_{\textrm{dir}}(\v m)\cdot\hat\rho_q(\v m)\\
%   \langle\Delta\v F_{\textrm{rec}}\rangle^\wedge(\v m)
%   &= -q_i\,\hat{\v K}^c_{\textrm{rec}}(\v m)\cdot\hat\rho_q(\v m)
% \end{align}
% If the charge distribution is \emph{locally} neutral, then all modes
% of the first order density distribution vanishes, and $\langle\Delta\v
% F_{\textrm{dir}}\rangle = \langle\Delta\v F_{\textrm{rec}}\rangle = 0$.

@@

% The naive calculation of the reciprocal Ewald summation
% \eqref{eqn:tmp6} will result in a computational cost of $\mathcal
% O(N^2)$. It has been shown that with a careful tuning of parameter
% $\beta$, real space cut-off radius $r_c$
% and the reciprocal space truncation $K_\alpha$,

@@

The SPME method provides two possibilities of calculating the
reciprocal force. The first one is to differentiate (with respect to
the particle position) the reciprocal term of the truncated Ewald
energy and then approximate the derived force \eqref{eqn:tmp8} by the
B-spline interpolation. This way is called the
\emph{ik-differentiation}.
The alternative way notices the high-order
smoothness of the B-spline interpolation, and derive the force by
differentiating the B-spline approximated reciprocal Ewald energy
(also truncated). The second way was proposed by the original SPME paper
\cite{essmann1995spm}, and is called the \emph{analytical
  differentiation}.
It has been shown that using the same parameters,
the ik-differentiation is more precise than the analytical
differentiation, but it uses twice more FFTs, which is a bottleneck
for the communication in parallel
implementations~\cite{wang2010optimizing}.
For the ik-differentiation, the calculated force is not the negative
gradient of the calculated energy. The analytical differentiation
does not have this problem, but it (to the approximation precision)
violates the Newton's third law.
Since the SPME method is
not the point of the present paper, we refer the readers to, for
example, Ref. \cite{essmann1995spm, deserno1998mue1,
  wang2010optimizing} for details of this method.

@@

% the last term in Eqn.~\eqref{eqn:ik.error.force.1},

@@

in the form of~\eqref{eqn:cor-kernel}, so the error estimate derived by
corollary~\ref{thm:tmp2} cannot be used here. However, it is still

@@

to the Appendix~\ref{sec:appendix}.  We notice the

@@

in Eqn.~\eqref{eqn:error-A} introduces some high-wave-number terms in

>>>>

\documentclass[journal=jacsat,manuscript=article]{achemso}
% \documentclass[aps,pre,preprint,unsortedaddress]{revtex4}
% \documentclass[aps,jcp,unsortedaddress]{revtex4-1}
%\documentclass[aps,pre,twocolumn]{revtex4-1}
%\documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}

\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage[dvips]{graphicx}
\usepackage{color}
\usepackage{tabularx}

@@

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
% \newcommand{\recheck}[1]{{\textbf{#1}}}
% \newcommand{\redc}[1]{{\textbf{#1}}}
\renewcommand{\v}[1]{\textbf{\textit{#1}}}
\renewcommand{\d}[1]{\textsf{#1}}
% \allowdisplaybreaks

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

@@

\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Pingwen Zhang}
\affiliation{LMAM and School of Mathematical Sciences, Peking University}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

\title[\texttt{achemso} demonstration]
{The Numerical Accuracy of Computing Electrostatic Interaction
  in the Inhomogeneous and Correlated Molecular Systems:
  for the Ewald Summation, SPME and Staggered Mesh Ewald Methods.
}

@@

@@

The aforementioned fast algorithms have achieved great success
over the past three decades.
However, the fact that the user should provide six working parameters
for these methods may be problematic: it is difficult to find the
most efficient parameter set in such a 6-dimensional parameter space.
% The experiential choices may be either less accurate or wasting
% computational effort.
% This problem can be solved by an
\recheck{
  This problem can be solved by comparing
  the forces calculated by
  using different parameter sets
  on a representative snapshot of the system,
  see e.g. Ref.~\cite{abraham2011optimization}.}
\recheck{An other approach is the 
  in-depth theoretical study}
of the error introduced by the algorithms, namely the
error estimate~\cite{kolafa1992cutoff, hummer1995numerical,
  petersen1995accuracy, deserno1998mue2, stern2008mesh,
  wang2010optimizing, neelov2010interlaced, ballenegger2012convert},
by which the error is described by a function of the working
parameters.  Using the error estimates, a work flow has been designed to
automatically determine the nearly optimal combination of parameters,
in the sense that it is the fastest one satisfying a demanding
accuracy~\cite{wang2010optimizing}. However, so far, all error
estimates assume the homogeneity of the charge distribution, and the
independency of any pair of charges (i.e. the charges are
uncorrelated), but neither of the assumptions is satisfactorily
fulfilled in most molecular
systems of practical interest.  For example, the lipid bilayer membrane
solving in water: the hydrophobic tail(s) of the lipids carry no charge, and
the charges in the system
are correlated by the covalent bonds, the van der Waals
interaction, the hydrogen bonding network, etc..  Therefore, the current
error estimates may be problematic, or even not applicable.  Several
papers have shown that due to the correlation of charges, the error estimates
obviously deviate from the real
error~\cite{deserno1998mue2,wang2010optimizing,ballenegger2012convert}.

The purpose of this manuscript is to develop a reliable  error
estimate in inhomogeneous and correlated molecular systems.
We estimate the force error because the application is the molecular
dynamics (MD) simulation.  Firstly, a general error estimate
framework for the pairwise interaction is set up.
\recheck{We show that the
  error may be approximately
  decomposed into three additive parts: the homogeneity error,
the inhomogeneity error and the correlation error~\cite{wang2012}.}
% These error estimates are only of theoretical interest, and
% not applicable, because the computational costs of the error estimates
% are $\mathcal O(N^2)$ (homogeneity and inhomogeneity error)
% and $\mathcal O(N^3)$ (correlation error).
% If the error force has a special form so that the error estimates
% can be written as convolutions, the computational costs
% are reduced to $\mathcal O(N\log N)$ and 
% $\mathcal O(N^2\log N)$, respectively.
The computational cost of the homogeneity and inhomogeneity
error is $\mathcal O(N\log N)$ when the error force kernel is
of convolution form.
The  correlation error can also be
estimated at a cost of $\mathcal O(N\log N)$
by the nearest neighbor approximation
technique, which \redc{assumes} the error  is dominated by  short-range correlation.
This
is the case for most the liquid and gas systems far from critical point.
Generally speaking,
when the error force can be written as a pairwise form, 
all estimate are derived easily under this framework.
Following this framework, we
develop the error estimates for the Ewald summation, the SPME 
and the staggered method Ewald methods.   The estimates
of the homogeneity and inhomogeneity errors are tested by two
randomly distributed inhomogeneous charges systems,
% (i.e. the charges are not correlated),
and is proved to
be very sharp.
The error estimates are also tested in a real inhomogeneous water system,
in which the charge correlation plays a very important role
on the force error. By 
including the nearest neighbor approximation of
\redc{the} correlation error, the quality of the estimates
is impressively enhanced.
This paper is ended by conclusions and remarks in the last
section.

@@

Here are some remarks concerning \ref{thm:tmp1} and Corollary 2:
%~\ref{thm:tmp2}:

@@

  \ref{eqn:def-rhoq} and \ref{eqn:def-rhoq2} are called \emph{the

@@

  $\rho^{(2)}(\v r, \v r')$ in \ref{eqn:def-rho2}

@@

  \ref{eqn:def-c} is the \emph{pairwise charge correlation function},

@@

  Also see the Example 3 presented later in this paper.

@@

  of the interaction (see \ref{eqn:thm-error-force}) is the key to derive
  the error estimates \ref{eqn:thm-meanf} and \ref{eqn:thm-error}.

@@

  By setting $q_i = q_j = 1$, the
  % corollary~\ref{thm:tmp2}
  Corollary~2
  proves

@@

  three terms on the r.h.s. of \ref{eqn:thm-error} are called the

@@

  The first term on the r.h.s. of \ref{eqn:error-split} is called

@@

  \ref{thm:tmp1} calculates the homogeneity and inhomogeneity

@@

  the
  % corollary~\ref{thm:tmp2},
  Corollary~2,

@@

\item A non-vanishing mean error force~\ref{eqn:thm-meanf} was proved

@@

Starting from \ref{eqn:prf-error-step3},

@@

\ref{eqn:def-rho2}.

@@

second \redc{term} on the r.h.s. of the \ref{eqn:split-rho-q2} is

@@

on the r.h.s. of \ref{eqn:split-rho-q2} is actually the

@@

By inserting \ref{eqn:fnc-correlation-water} into

@@

As an example, we consider a typical three point charge water model
TIP3P~\cite{jorgensen1983comparison}: each hydrogen atom carries a
charge of $q_{\textrm{H}} = +0.417e$ and each oxygen atom carries a
charge of $q_{\textrm{O}} = -0.834e$.  \recheck{ There are two types
  of correlation in this system: (1) the bonded correlation
  that originates from the rigid H-O
  bond and H-O-H angle within one molecule,
  and (2) the
  non-bonded correlation or the
  intermolecular correlation that originates from
  a complex coupling of various intermolecular interactions, such as
  the van der Waals interaction,
  the hydrogen bonding and etc..
% Within one molecule, the H-O bond and
% H-O-H angle are rigid, so the correlation within one molecule is
% considered important.
% Therefore, in this case,
}

@@

neighbors by the atoms within the same water molecule.  Denoting the

@@

The two ensemble averages in \ref{eqn:c-error-1}

@@

\ref{eqn:c-error-1} is by the Fourier transform

@@

the \ref{eqn:c-error-1} becomes:

@@

system, as illustrated by \ref{fig:tmp-add0}.

@@

% estimate~\ref{eqn:c-error-water}

@@

standard long-range calculations.}

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{correlation.eps}
  \caption{The schematic plot of two interacting water molecules.
    When we calculate the force error of the left oxygen atom (it is now
    a testing charge), the
    correlation with its bonded hydrogens are not considered.
    The correlations within any neighboring molecule are considered,
    as indicated by the solid lines with two arrows.}
  \label{fig:tmp-add0}
\end{figure}

\recheck{
  For the non-bonded correlation,
  we define the nearest neighbors
  by non-bonded atoms closer than a certain radius $R_g$ (neighboring range).
  If the system is homogeneous and isotropic, the pair
  correlation function $\rho^{(2)}(\v r', \v r'')$ in
  \ref{eqn:start-corr} can be represented by the radial
  distribution functions (RDFs).
  Following the same way as the bonded correlation,
  the nearest neighbor approximation of
  the non-bonded correlation error is estimated:
  the functions $T_{\textrm{O}}$ and $T_{\textrm{H}}$
  are given by
  \begin{align}\label{eqn:nb-nna-1}
    \hat T_{\textrm{O}} (\v m)
    &= \int_0^{R_g}
    [q_{_{\textrm{H}}} \rho_{_{\textrm{H}}} \tilde g_{_{\textrm{OH}}}(r)
    +q_{_{\textrm{O}}} \rho_{_{\textrm{O}}} \tilde g_{_{\textrm{OO}}}(r)]
    \,\frac{\sin(2\pi m r)}{2\pi mr}\,4\pi r^2 \d dr,\\\label{eqn:nb-nna-2}
    \hat T_{\textrm{H}} (\v m)
    &= \int_0^{R_g}
    [q_{_{\textrm{H}}} \rho_{_{\textrm{H}}} \tilde g_{_{\textrm{HH}}}(r)
    +q_{_{\textrm{O}}} \rho_{_{\textrm{O}}} \tilde g_{_{\textrm{OH}}}(r)]
    \,\frac{\sin(2\pi m r)}{2\pi mr}\,4\pi r^2 \d dr,
  \end{align}
  where
  $\tilde g_{_{\textrm{OO}}}(r) = g_{_{\textrm{OO}}}(r) - 1$,
  $\tilde g_{_{\textrm{OH}}}(r) = g_{_{\textrm{OH}}}(r) - 1$ and
  $\tilde g_{_{\textrm{HH}}}(r) = g_{_{\textrm{HH}}}(r) - 1$.
  $g_{_{\textrm{OO}}}(r)$, $g_{_{\textrm{OH}}}(r)$ and $g_{_{\textrm{HH}}}(r)$
  are the RDFs of oxygen-oxygen, oxygen-hydrogen and hydrogen-hydrogen,
  respectively. $\rho_{_{\textrm{H}}}$ and $\rho_{_{\textrm{O}}}$ are
  the homogeneous number densities
  of the hydrogen and oxygen atoms, respectively.
  For a short-range correlated system, the RDFs converges to 1
  as $r$ goes to infinity, so the integrals in \ref{eqn:nb-nna-1}
  and \ref{eqn:nb-nna-2} are convergent with respect to
  increasing $R_g$.

@@

The complementary error function $\textrm{erfc}(r)$ in
\ref{eqn:tmp2} decays exponentially as $r$ goes to infinity.
Therefore, it is a short-range interaction in the real space, which
can be calculated by the standard cut-off and neighbor list
method~\cite{frenkel02b} at a cost of $\mathcal O(N)$.  The reciprocal
part also decays exponentially as the magnitude of the Fourier mode
$\vert\v m\vert$ increases. Therefore, in practice, the infinite
summation in \ref{eqn:tmp3} is truncated, and only a finite
summation is calculated:
% Both the direct and the reciprocal parts are calculated by cut-off in
% real and reciprocal space, respectively.
% For the reciprocal part the cut-offed energy reads
\begin{align}\label{eqn:tmp6}
  E^{\textrm{tr}}_{\textrm{rec}} & =
  % \frac1{2\pi V}
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  % \frac{\exp(-\pi^2\v m^2 / \beta^2)}{\v m^2} S(\v m) S(-\v m)   
  f(\v m)\, S(\v m) S(-\v m),
\end{align}
where, for short,
\begin{align}
  f(\v m) := \frac{\exp(-\pi^2 \v m^2 / \beta^2)}{2\pi V \v m^2},
\end{align}
and $K_\alpha$ is the number of Fourier modes used on direction
$\alpha$.  The truncated reciprocal force acting on particle $i$ is
\begin{align}
  % \nonumber
  \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
  % &= 
  % q_i 
  % \sum_{
  %   \begin{subarray}{c}
  %     \vert m_\alpha\vert < K_\alpha/2\\
  %     \v m\neq 0
  %   \end{subarray}}
  % \v g(\v m) \,
  % S(-\v m)\,
  % e^{2\pi i\v m\cdot\v r_i}, \\
  \label{eqn:tmp8}
  &= 
  \sum_j   q_iq_j
  \sum_{
    \begin{subarray}{c}
      \vert m_\alpha\vert < K_\alpha/2\\
      \v m\neq 0
    \end{subarray}}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r_{ij}},
\end{align}
where
\begin{align}
  \v g(\v m) = -4 \pi \v m i f(\v m).
\end{align}

@@

Both the error force kernel of the direct and reciprocal
part of the Ewald summation can be
written in the form of \ref{eqn:cor-kernel}.  The direct
part is:
% The direct part is a short-range interaction in real space, so it is
% usually treated by the cut-off method, which gives a error force
% kernel of
\begin{align}
  \v K^{\textrm{cut}}_{\textrm{dir}}(\v r) =
  \left\{
    \begin{aligned}
      &\,0, &\quad & r \leq r_c;\\
      &
      \Big[
      \frac{2\beta}{\sqrt\pi} e^{-\beta^2r^2} + \frac{\textrm{erfc}(\beta r)}{r}
      \Big]\frac{\v r}{r^2}
      ,& & r > r_c,
      % \frac{\textrm{erfc}(\beta\vert\v r\vert)}{\vert\v r\vert}
    \end{aligned}
  \right.
\end{align}
where $r = \vert\v r\vert$, and $r_c$ is the cut-off radius in the
real space.
% The reciprocal error force, namely the difference between
% the the exact and truncated reciprocal force, is
% \begin{align}
%   \Delta \v F^{\textrm{tr}}_{\textrm{rec}}(\v r_i)
%   = & - 
%   \sum_j\,q_iq_j
%   \sum_{\vert m_\alpha\vert \geq K_\alpha/2}
%   \v g(\v m) \,
%   e^{2\pi i\v m\cdot\v r_{ij}}.
% \end{align}
% Therefore,
The reciprocal error force kernel is given by:
\begin{align}
  \v K^{\textrm{tr}}_{\textrm{rec}}(\v r) =
  \sum_{
      \vert m_\alpha\vert \geq K_\alpha/2}
  \v g(\v m) \,
  e^{2\pi i\v m\cdot\v r}.
\end{align}
By the
% corollary~\ref{thm:tmp2},
Corollary~2,
the error estimate of the
truncated Ewald summation is straightforward.

% Ref. \cite{short} pointed out that the inhomogeneity error is more
% harmful than the homogeneity error, so it is interesting to analyze
% the inhomogeneity error. Apply the Fourier transform on both sides of
% \ref{eqn:tmp10} and \ref{eqn:tmp18}:
% \begin{align}  
%   \langle\Delta\v F_{\textrm{dir}}\rangle^\wedge(\v m)
%   &= q_i\,\hat{\v K}^c_{\textrm{dir}}(\v m)\cdot\hat\rho_q(\v m)\\
%   \langle\Delta\v F_{\textrm{rec}}\rangle^\wedge(\v m)
%   &= -q_i\,\hat{\v K}^c_{\textrm{rec}}(\v m)\cdot\hat\rho_q(\v m)
% \end{align}
% If the charge distribution is \emph{locally} neutral, then all modes
% of the first order density distribution vanishes, and $\langle\Delta\v
% F_{\textrm{dir}}\rangle = \langle\Delta\v F_{\textrm{rec}}\rangle = 0$.

@@

% The naive calculation of the reciprocal Ewald summation
% \ref{eqn:tmp6} will result in a computational cost of $\mathcal
% O(N^2)$. It has been shown that with a careful tuning of parameter
% $\beta$, real space cut-off radius $r_c$
% and the reciprocal space truncation $K_\alpha$,

@@

The SPME method provides two possibilities of calculating the
reciprocal force. The first one is to differentiate (with respect to
the particle position) the reciprocal term of the truncated Ewald
energy and then approximate the derived force \ref{eqn:tmp8} by the
B-spline interpolation. This way is called the
\emph{ik-differentiation}.
The alternative way notices the high-order
smoothness of the B-spline interpolation, and derive the force by
differentiating the B-spline approximated reciprocal Ewald energy
(also truncated). The second way was proposed by the original SPME paper
\cite{essmann1995spm}, and is called the \emph{analytical
  differentiation}.
It has been shown that using the same parameters,
the ik-differentiation is more precise than the analytical
differentiation, but it uses twice more FFTs, which is a bottleneck
for the communication in parallel
implementations~\cite{wang2010optimizing}.
For the ik-differentiation, the calculated force is not the negative
gradient of the calculated energy. The analytical differentiation
does not have this problem, but it (to the approximation precision)
violates the Newton's third law.
Since the SPME method is
not the point of the present paper, we refer the readers to, for
example, Ref. \cite{essmann1995spm, deserno1998mue1,
  wang2010optimizing} for details of this method.

@@

% the last term in \ref{eqn:ik.error.force.1},

@@

in the form of~\ref{eqn:cor-kernel}, so the error estimate derived by
% corollary~\ref{thm:tmp2}
Corollary~2
cannot be used here. However, it is still

@@

to the appendix.  We notice the

@@

in \ref{eqn:error-A} introduces some high-wave-number terms in
